Forecasting the 2024 U.S. Presidential Election\thanks{Code and data are available at: {[}https://github.com/AndyYanxunJiang/president-polls).}

### 2024-11-04 paper/paper.tex

---
title: "Forecasting the 2024 U.S. Presidential Election"
subtitle: "Using Aggregated Polling Data to Predict the Outcome of the 2024 Presidential Race"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/president-polls)."
date: today
date-format: long
abstract: "The 2024 U.S. presidential election is shaping up to be a highly swing race, with polling data playing a crucial role in gauging public sentiment and forecasting the potential outcome. This paper aims to build a predictive model using a poll-of-polls approach, aggregating data from multiple pollsters to forecast the winner of the upcoming election. The analysis incorporates polling data from various sources, filtered to include only high-quality pollsters with a numeric grade above 3, to ensure the reliability of the predictions. Additionally, a detailed examination of one selected pollster’s methodology is conducted, highlighting strengths, weaknesses, and biases in the survey process. The paper concludes with an idealized polling methodology, designed within a $100K budget, to enhance the accuracy of election forecasts. Results are presented in the context of understanding polling trends, potential biases, and the evolving landscape of voter preferences as the election approaches."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(readr)
library(here)
library(knitr)
library(kableExtra)
library(scales)
library(arrow)
library(bayesplot)
library(rstanarm)
data_filtered <- read.csv(here::here("data/02-analysis_data/model_data.csv"))
data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
```

# Introduction
The 2024 U.S. presidential election is shaping up to be one of the most closely watched and potentially pivotal elections in recent history. As public opinion shifts and political dynamics evolve, polling data has emerged as a critical tool for forecasting the outcome and understanding voter preferences. Polls provide a snapshot of where candidates stand at various points in time, reflecting the impact of campaigns, debates, and major events. However, the reliability and accuracy of individual polls can vary significantly based on methodology, sample size, and other factors, making it necessary to aggregate multiple sources of data for a more robust prediction.

This paper aims to forecast the outcome of the 2024 U.S. presidential election by leveraging a "poll-of-polls" approach, which aggregates data from various pollsters to create a more comprehensive picture of the race. By focusing on high-quality polls and conducting a detailed analysis of the polling trends for the two main candidates, Donald Trump and Kamala Harris, we seek to uncover patterns in voter support and potential factors influencing shifts in public opinion. Additionally, the paper examines the methodologies of different pollsters to assess their strengths and weaknesses, which can help identify possible biases in the aggregated data.

The analysis will further explore the geographic distribution of support across key swing states, providing insights into the electoral landscape. Finally, the paper presents an idealized polling methodology with a $100,000 budget, designed to improve the accuracy and representativeness of election forecasts. By integrating multiple data sources and critically evaluating polling practices, this study aims to contribute to the ongoing discussion on the role of polling in democratic processes and election forecasting.

# Data {#sec-data}
The data used in this paper was obtained from the FiveThirtyEight website [@Best2024], which aggregates polling data and other relevant indicators to track the U.S. election landscape. We accessed this data using custom web scraping tools and APIs provided by FiveThirtyEight. The dataset includes information on pollster, state, start and end dates of polls, sample size, population type (e.g., registered voters or likely voters), and party affiliation.

## Overview
Our analysis was conducted using R [@citeR] along with several supporting packages. We used the **tidyverse** for data manipulation [@tidyverse], **readr** for data import [@readr], and **here** for file management [@here].

Dynamic reporting was handled with **knitr** [@knitr], while **kableExtra** allowed for enhanced table formatting [@kableExtra]. We used **scales** to adjust visualization scales [@scales], and **arrow** for efficient data processing [@arrow].

For Bayesian modeling and visualization, we used **bayesplot** [@bayesplot] and **rstanarm** [@rstanarm].

## Measurement
The FiveThirtyEight dataset is updated in real-time, providing the latest polling data for the 2024 U.S. presidential election. This dataset aggregates information from multiple reputable polling sources, applying adjustments to enhance reliability, such as weighting based on pollster quality and recency of data. These measures ensure that the dataset remains a dependable reflection of current voter sentiment and trends as Election Day approaches.

## Analysis Data
The analysis dataset includes several key variables: State, Sample Size, Population Type (e.g., Registered Voters), Party Affiliation, Candidate Preference (Trump or Harris), and Pollster. These variables allow for an analysis of voter sentiment in specific swing states, with a focus on support levels between Donald Trump and Kamala Harris.

Certain variables, such as Question ID, Start Date, End Date, and Hypothetical Questions, were removed as they were not necessary for this analysis. Additionally, data on other candidates was excluded to concentrate solely on the Trump vs. Harris matchup, aligning with the focus on the most relevant contenders in the seven swing states: Pennsylvania, Georgia, North Carolina, Michigan, Arizona, Wisconsin, Nevada.

@tbl-sample provides a random sample of the analysis dataset, displaying all relevant variables used for analysis.

```{r}
#| label: tbl-sample
#| tbl-cap: Random sample of 10 rows from the analysis data.
#| echo: false
#| warning: false
#| message: false

set.seed(311)

# Sample 10 random observations from the filtered data with Trump and Harris, and remove unnecessary columns
sample_data <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  sample_n(10) %>%
  select(-start_date, -end_date, -question_id, -hypothetical, -pct)

# Display the sampled data as a table with refined styling
kable(sample_data, caption = "Sample of 10 Random Observations from the Analysis Data") %>%
  kable_styling(full_width = FALSE, position = "center") %>%  # Center the table in the PDF
  column_spec(1, width = "3cm") %>%    # Adjust specific columns for better fit
  column_spec(2, width = "2.5cm") %>%
  column_spec(3, extra_css = "padding-left: -5px;") %>%  # Move sample_size column slightly to the left
  column_spec(4:5, width = "1.5cm")    # Adjust remaining columns with less text

```

## Predictor variables
End date numeric:
This variable represents the numeric conversion of the poll end date, indicating the time at which the polling data was collected. By transforming dates into numeric values, we can account for temporal trends and shifts in voter sentiment. This predictor helps identify changes in support over time and captures any momentum shifts that occur as the election draws closer.

Candidate name:
This categorical variable distinguishes between the main candidates: Donald Trump and Kamala Harris. By including candidate-specific factors, the model can account for differences in support levels attributed to individual characteristics or public perception of each candidate. Additionally, this variable enables the model to gauge the relative standing between the two contenders in different states and overall.

# Model
In this analysis, we utilize a logistic regression model to predict polling percentages for the 2024 U.S. presidential candidates over time. In order to focus on states with historical significance in elections. The model is specified as a Generalized Linear Model (GLM) with a binomial family and a logit link function, appropriate for modeling binary or proportion-based outcomes.

## Model set-up
The dataset, model_data.csv, provides polling data with average_pct as the observed average polling percentage among 7 swing states. The end_date indicates the polling period's endpoint.
To predict polling percentages beyond the available data, we created an additional sequence of 20 days beyond the dataset's last recorded date. Because we want to predict the result in November 5. This expansion allows for forecasting near-term polling trends. Also, capturing potential shifts in candidate support leading up to election day. The model was trained on existing data and applied to this extended date range using predict(model, newdata = ., type = "response").

Our model setup and variable selection are as follows:
$$
\text{logit}(p_i) = \beta_0 + \beta_1 (\text{Date}) + \beta_2 (\text{Candidate})
$$

where:
- $$p_i \ \text{represents the probability of support for candidate i at a specific date.}$$
- $$\beta_0 \ \text{is the intercept, representing baseline support.}$$
- $$\beta_1 \ \text{captures changes in support over time (date effect).}$$
- $$\beta_2 \ \text{is a categorical variable accounting for each candidate.}$$

### Model justification
The choice of a logistic regression model was guided by the nature of polling data. Nature of polling data represents proportions that naturally bound between 0 and 1. The logistic function’s S-shaped curve effectively models the relationship between time and polling percentage, accommodating non-linear trends that may arise as election day nears.
Firstly, candidate names should be included as a categorical predictor in the model. It allows us to capture individual impact in support rates by candidate. In addition, a transformation from date to numeric is made on end_date in order to get a numerical format. It is called the end_date_numeric. It helps put date as a predictor in the model. By structuring the model this way, it can predict the percentage of the poll by both the impact of candidates and trends over time.

# Results
The analysis of polling data across key swing states for the 2024 U.S. Presidential Election reveals insights into the levels of support for candidates Donald Trump and Kamala Harris. The findings are presented through three primary visualizations: the average polling percentage by state, the distribution of polling percentages, and the difference in average polling percentage between the two candidates.

## Average Polling Percentage by State
From @fig-1, we can see that in the swing states of Arizona, Georgia, and North Carolina, Trump holds a narrow lead. In contrast, Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When looking at the overall average across these states, Harris slightly edges out Trump, indicating her overall advantage in these swing areas. In our assumed electoral map, this projects Kamala Harris over the edge with 276 electoral votes as opposed to Donald Trump's 262. This shows that our model predicts Kamala Harris to win the 2024 Presidential Election if she is able to maintain her lead over all of Pennsylvania, Wisconsin, Nevada, and Michigan.

```{r}
#| label: fig-1
#| fig-cap: "Average Polling Percentage by State"
#| echo: false
#| warning: false
#| message: false
# Calculate the average polling percentage by state for Trump and Harris
data_avg <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state, candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE), .groups = "drop")

# Calculate the overall average polling percentage for Trump and Harris
overall_avg <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE)) %>%
  mutate(state = " Overall")  # Add a space before "Overall" to separate it visually

# Combine state-level data with overall average
combined_data <- bind_rows(data_avg, overall_avg)

# Ensure "Overall" appears at the bottom with spacing
combined_data$state <- factor(combined_data$state, levels = c(setdiff(unique(combined_data$state), " Overall"), " Overall"))

# Plot the average polling percentages by state with overall comparison
ggplot(combined_data, aes(x = state, y = avg_pct, fill = candidate_name)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), aes(color = ifelse(state == " Overall", "Overall", "State")), width = 0.6) + 
  scale_color_manual(values = c("Overall" = "black", "State" = NA), guide = "none") +  # Add dark grey outline for "Overall" bars
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +  # Custom colors for consistency
  labs(title = "Average Polling Percentages by State",
       x = "State",
       y = "Average Polling Percentage",
       fill = "Candidate") +
  coord_flip() +  # Flip coordinates for easier reading
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.y = element_text(margin = margin(r = 10)), # Add spacing to y-axis labels
    panel.grid.major.y = element_blank(), # Remove y-axis gridlines for clarity
    panel.spacing = unit(1, "lines") # Increase panel spacing
  )

```
## Polling Percentage Distribution
The box plot in @fig-2 demonstrates the range and consistency of support for both candidates across each state. Trump shows a wider range of support, reflecting variability, while Harris has relatively steady polling percentages across states, maintaining higher consistency. Harris’s support is less variable and generally centers around higher polling percentages compared to Trump’s.

```{r}
#| label: fig-2
#| fig-cap: "Polling Percentage Distribution"
#| echo: false
#| warning: false
#| message: false
ggplot(data %>% filter(candidate_name %in% c("Donald Trump", "Kamala Harris")), 
       aes(x = state, y = pct, fill = candidate_name)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Polling Percentage Distribution by State",
       x = "State",
       y = "Polling Percentage",
       fill = "Candidate") +
  scale_fill_manual(values = c("Donald Trump" = "#F8766D", "Kamala Harris" = "#56B4E9")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
## Difference in Average Polling Percentage
@fig-3 highlights the difference in average polling percentages by state, with positive values indicating Trump’s lead and negative values indicating Harris's. Trump’s lead in the three states (Pennsylvania, Wisconsin, Nevada, and Michigan) is relatively narrow, with margins of less than 1 percent. In contrast, Harris’s lead in the states favoring her (Pennsylvania, Wisconsin, Nevada, and Michigan) ranges from 1 to 4 percent, indicating her stronger support in the states where she leads. Consequently, this difference in average polling percentages show that Kamala Harris' lead of 276 electoral votes is supported by a reliable margin and Donald Trump's lead in Arizona, Georgia, and North Carolina is turbulent.

```{r}
#| label: fig-3
#| fig-cap: "Difference in Average Polling Percentage"
#| echo: false
#| warning: false
#| message: false
data_diff <- data %>%
  filter(candidate_name %in% c("Donald Trump", "Kamala Harris")) %>%
  group_by(state) %>%
  summarize(diff_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE) - 
                      mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE), .groups = "drop")

ggplot(data_diff, aes(x = reorder(state, diff_pct), y = diff_pct, fill = diff_pct > 0)) +
  geom_bar(stat = "identity", color = "darkgrey") +
  scale_fill_manual(values = c("TRUE" = "#F8766D", "FALSE" = "#56B4E9"), 
                    labels = c("Kamala Harris", "Donald Trump")) +
  labs(title = "Difference in Average Polling Percentage by State",
       x = "State",
       y = "Difference in Polling Percentage",
       fill = "Lead") +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Swing State Polling Prediction Model

@fig-4 shows the trend of support rate changes for each candidate before the upcoming election, as well as the model's predicted values for the next 20 days.We predicts that Kamala Harris may win based on our model.
However, this shows the trend of support rate changes for each candidate before the upcoming election, as well as the model's predicted values for the days until the election. This support rate changes model exhibits a higher rate of growth in Donald Trump's share of support. Although Kamala Harris is in the lead, Donald Trump's growth of support rate is an element that could stifle Harris' lead and drop her below her required 270 electoral votes.

```{r}
#| label: fig-4
#| fig-cap: "Swing State Polling Prediction Model"
#| echo: false
#| warning: false
#| message: false

invisible(Sys.setlocale("LC_TIME", "en_US"))
model_data <- read_csv(here::here("data/02-analysis_data/model_data.csv"))
model <- readRDS(here::here("models/model.rds"))

model_data <- model_data %>%
  mutate(predicted_pct = predict(model, newdata = ., type = "response"))
last_date <- max(model_data$end_date, na.rm = TRUE)
new_dates <- seq.Date(from = last_date + 1, by = "day", length.out = 20)
new_data <- expand.grid(
  end_date = new_dates,
  candidate_name = unique(model_data$candidate_name)
)
new_data <- new_data %>%
  mutate(
    end_date_numeric = as.numeric(end_date), 
  ) %>% mutate(predicted_pct = predict(model, newdata = ., type = "response"))
all_data <- bind_rows(model_data, new_data)

ggplot(all_data, aes(x = end_date, y = pct, color = candidate_name)) +
  geom_point(aes(y = average_pct/100), alpha = 0.5) + 
  geom_line(aes(y = predicted_pct), size = 1) + 
  labs(title = "2024 US President Polls - Predicted vs True",
       x = "Date",
       y = "Rate of support") +
  theme_minimal() +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", 
               limits = as.Date(c("2024-01-01", "2024-11-5"))) +
  scale_y_continuous(labels = percent_format(), limits = c(0.44, 0.47))

```

# Discussion
## Turbulent Margins
Based on our results from @fig-1 predicted that Kamala Harris to win the presidential election with 276 electoral votes. Furthermore, @fig-2 showed that her lead in the states of Pennsylvania, Wisconsin, Nevada, and Michigan are stronger than Donald Trump's support in Pennsylvania, Wisconsin, Nevada, and Michigan. From @fig-2, we can also observe that Kamala Harris' lead in Nevada seems to be the weakest. In the framework of the electoral college system, losing Nevada to Donald Trump would not change the results as Nevada holds 6 electoral votes, which means that Kamala Harris would still possess the necessary 270 to win the election. However, this is not the case if Kamala Harris loses support in any of the three other states of Pennsylvania, Wisconsin, and Michigan as this would place her beneath the 270 seat threshold. In fact, losing any one of these states would result in Donald Trump's victory. @fig-3 showed that the support rate of Donald Trump is increasing faster than that of Kamala Harris', threatening Harris' lead over Pennsylvania, Wisconsin, and Michigan.

## Impact of Key Demographics and Policy Appeals
Harris’s lead in several swing states may be influenced by her campaign’s appeal to key demographic groups, particularly immigrants. The Democratic platform has emphasized social and economic benefits for immigrant communities, which may have strengthened her base in states with significant immigrant populations. Policies that promise support for healthcare, education, and other public services are likely contributing to Harris’s appeal among these groups, solidifying her lead in states like Nevada and Michigan. This demographic support provides Harris with a strategic advantage, as these communities may play a crucial role in voter turnout and loyalty.

## Influence of Endorsements and High-Profile Backing
Recently, high-profile figures like Elon Musk have publicly thrown their support behind Trump, leveraging their influence to rally support for his campaign [@Zahn2024]. Musk’s endorsement could invigorate Trump’s base, especially among Musk’s followers and those engaged in the fields of business and technology. This visible backing from such influential figures has the potential to sway undecided voters and influence last-minute decisions, particularly in swing states where Trump’s lead remains narrow.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting the 2024 U.S. Presidential Election},
  pdfauthor={Andy Jiang; Onon Burentuvshin; Rui Hu},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Forecasting the 2024 U.S. Presidential Election\thanks{Code and
data are available at:
{[}https://github.com/AndyYanxunJiang/president-polls).}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Using Aggregated Polling Data to Predict the Outcome of the
2024 Presidential Race}
\author{Andy Jiang \and Onon Burentuvshin \and Rui Hu}
\date{November 4, 2024}

\begin{document}
\maketitle
\begin{abstract}
The 2024 U.S. presidential election is shaping up to be a highly swing
race, with polling data playing a crucial role in gauging public
sentiment and forecasting the potential outcome. This paper aims to
build a predictive model using a poll-of-polls approach, aggregating
data from multiple pollsters to forecast the winner of the upcoming
election. The analysis incorporates polling data from various sources,
filtered to include only high-quality pollsters with a numeric grade
above 3, to ensure the reliability of the predictions. Additionally, a
detailed examination of one selected pollster's methodology is
conducted, highlighting strengths, weaknesses, and biases in the survey
process. The paper concludes with an idealized polling methodology,
designed within a \$100K budget, to enhance the accuracy of election
forecasts. Results are presented in the context of understanding polling
trends, potential biases, and the evolving landscape of voter
preferences as the election approaches.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The 2024 U.S. presidential election is shaping up to be one of the most
closely watched and potentially pivotal elections in recent history. As
public opinion shifts and political dynamics evolve, polling data has
emerged as a critical tool for forecasting the outcome and understanding
voter preferences. Polls provide a snapshot of where candidates stand at
various points in time, reflecting the impact of campaigns, debates, and
major events. However, the reliability and accuracy of individual polls
can vary significantly based on methodology, sample size, and other
factors, making it necessary to aggregate multiple sources of data for a
more robust prediction.

This paper aims to forecast the outcome of the 2024 U.S. presidential
election by leveraging a ``poll-of-polls'' approach, which aggregates
data from various pollsters to create a more comprehensive picture of
the race. By focusing on high-quality polls and conducting a detailed
analysis of the polling trends for the two main candidates, Donald Trump
and Kamala Harris, we seek to uncover patterns in voter support and
potential factors influencing shifts in public opinion. Additionally,
the paper examines the methodologies of different pollsters to assess
their strengths and weaknesses, which can help identify possible biases
in the aggregated data.

The analysis will further explore the geographic distribution of support
across key swing states, providing insights into the electoral
landscape. Finally, the paper presents an idealized polling methodology
with a \$100,000 budget, designed to improve the accuracy and
representativeness of election forecasts. By integrating multiple data
sources and critically evaluating polling practices, this study aims to
contribute to the ongoing discussion on the role of polling in
democratic processes and election forecasting.

\hypertarget{sec-data}{%
\section{Data}\label{sec-data}}

The data used in this paper was obtained from the FiveThirtyEight
website (Best, Bycoffe, et al. 2024), which aggregates polling data and
other relevant indicators to track the U.S. election landscape. We
accessed this data using custom web scraping tools and APIs provided by
FiveThirtyEight. The dataset includes information on pollster, state,
start and end dates of polls, sample size, population type (e.g.,
registered voters or likely voters), and party affiliation.

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

Our analysis was conducted using R (R Core Team 2022) along with several
supporting packages. We used the \textbf{tidyverse} for data
manipulation (Wickham et al. 2019), \textbf{readr} for data import
(Wickham, Hester, and Bryan 2023), and \textbf{here} for file management
(Müller 2020).

Dynamic reporting was handled with \textbf{knitr} (Xie 2024), while
\textbf{kableExtra} allowed for enhanced table formatting (Zhu 2024). We
used \textbf{scales} to adjust visualization scales (Wickham, Pedersen,
and Seidel 2023), and \textbf{arrow} for efficient data processing
(Richardson et al. 2024).

For Bayesian modeling and visualization, we used \textbf{bayesplot}
(Gabry and Mahr 2024) and \textbf{rstanarm} (Goodrich et al. 2024).

\hypertarget{measurement}{%
\subsection{Measurement}\label{measurement}}

The FiveThirtyEight dataset is updated in real-time, providing the
latest polling data for the 2024 U.S. presidential election. This
dataset aggregates information from multiple reputable polling sources,
applying adjustments to enhance reliability, such as weighting based on
pollster quality and recency of data. These measures ensure that the
dataset remains a dependable reflection of current voter sentiment and
trends as Election Day approaches.

\hypertarget{analysis-data}{%
\subsection{Analysis Data}\label{analysis-data}}

The analysis dataset includes several key variables: State, Sample Size,
Population Type (e.g., Registered Voters), Party Affiliation, Candidate
Preference (Trump or Harris), and Pollster. These variables allow for an
analysis of voter sentiment in specific swing states, with a focus on
support levels between Donald Trump and Kamala Harris.

Certain variables, such as Question ID, Start Date, End Date, and
Hypothetical Questions, were removed as they were not necessary for this
analysis. Additionally, data on other candidates was excluded to
concentrate solely on the Trump vs.~Harris matchup, aligning with the
focus on the most relevant contenders in the seven swing states:
Pennsylvania, Georgia, North Carolina, Michigan, Arizona, Wisconsin,
Nevada.

Table~\ref{tbl-sample} provides a random sample of the analysis dataset,
displaying all relevant variables used for analysis.

\hypertarget{tbl-sample}{}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{2.5cm}>{}r>{\raggedright\arraybackslash}p{1.5cm}>{\raggedright\arraybackslash}p{1.5cm}ll}
\caption{\label{tbl-sample}Random sample of 10 rows from the analysis data. }Sample of 10 Random Observations from the Analysis Data}\\
\toprule
pollster & state & sample\_size & population & party & answer & candidate\_name\\
\midrule
PPP & Arizona & 596 & rv & REP & Trump & Donald Trump\\
Morning Consult & Wisconsin & 700 & rv & REP & Trump & Donald Trump\\
Morning Consult & Georgia & 800 & rv & REP & Trump & Donald Trump\\
Morning Consult & Nevada & 574 & rv & REP & Trump & Donald Trump\\
Bullfinch & Pennsylvania & 800 & rv & DEM & Harris & Kamala Harris\\
\addlinespace
Public Opinion Strategies & Pennsylvania & 500 & rv & REP & Trump & Donald Trump\\
High Point University & North Carolina & 804 & rv & REP & Trump & Donald Trump\\
Emerson & Michigan & 1000 & rv & REP & Trump & Donald Trump\\
Quinnipiac & Georgia & 1203 & rv & REP & Trump & Donald Trump\\
Emerson & Pennsylvania & 850 & rv & REP & Trump & Donald Trump\\
\bottomrule
\end{longtable}

\hypertarget{predictor-variables}{%
\subsection{Predictor variables}\label{predictor-variables}}

\hypertarget{model}{%
\section{Model}\label{model}}

In this analysis, we utilize a logistic regression model to predict
polling percentages for the 2024 U.S. presidential candidates over time.
In order to focus on states with historical significance in elections.
The model is specified as a Generalized Linear Model (GLM) with a
binomial family and a logit link function, appropriate for modeling
binary or proportion-based outcomes.

\hypertarget{model-set-up}{%
\subsection{Model set-up}\label{model-set-up}}

The dataset, model\_data.csv, provides polling data with average\_pct as
the observed average polling percentage among 7 swing states. The
end\_date indicates the polling period's endpoint. To predict polling
percentages beyond the available data, we created an additional sequence
of 20 days beyond the dataset's last recorded date. Because we want to
predict the result in November 5. This expansion allows for forecasting
near-term polling trends. Also, capturing potential shifts in candidate
support leading up to election day. The model was trained on existing
data and applied to this extended date range using predict(model,
newdata = ., type = ``response'').

Our model setup and variable selection are as follows: \[
\text{logit}(p_i) = \beta_0 + \beta_1 (\text{Date}) + \beta_2 (\text{Candidate})
\]

where: -
\[p_i \ \text{represents the probability of support for candidate i at a specific date.}\]
- \[\beta_0 \ \text{is the intercept, representing baseline support.}\]
-
\[\beta_1 \ \text{captures changes in support over time (date effect).}\]
-
\[\beta_2 \ \text{is a categorical variable accounting for each candidate.}\]

\hypertarget{model-justification}{%
\subsubsection{Model justification}\label{model-justification}}

The choice of a logistic regression model was guided by the nature of
polling data. Nature of polling data represents proportions that
naturally bound between 0 and 1. The logistic function's S-shaped curve
effectively models the relationship between time and polling percentage,
accommodating non-linear trends that may arise as election day nears.
Firstly, candidate names should be included as a categorical predictor
in the model. It allows us to capture individual impact in support rates
by candidate. In addition, a transformation from date to numeric is made
on end\_date in order to get a numerical format. It is called the
end\_date\_numeric. It helps put date as a predictor in the model. By
structuring the model this way, it can predict the percentage of the
poll by both the impact of candidates and trends over time.

\hypertarget{results}{%
\section{Results}\label{results}}

The analysis of polling data across key swing states for the 2024 U.S.
Presidential Election reveals insights into the levels of support for
candidates Donald Trump and Kamala Harris. The findings are presented
through three primary visualizations: the average polling percentage by
state, the distribution of polling percentages, and the difference in
average polling percentage between the two candidates.

\hypertarget{average-polling-percentage-by-state}{%
\subsection{Average Polling Percentage by
State}\label{average-polling-percentage-by-state}}

From Figure~\ref{fig-1}, we can see that in the swing states of Arizona,
Georgia, and North Carolina, Trump holds a narrow lead. In contrast,
Harris leads in Pennsylvania, Wisconsin, Nevada, and Michigan. When
looking at the overall average across these states, Harris slightly
edges out Trump, indicating her overall advantage in these swing areas.

\hypertarget{polling-percentage-distribution}{%
\subsection{Polling Percentage
Distribution}\label{polling-percentage-distribution}}

The box plot in Figure~\ref{fig-2} demonstrates the range and
consistency of support for both candidates across each state. Trump
shows a wider range of support, reflecting variability, while Harris has
relatively steady polling percentages across states, maintaining higher
consistency. Harris's support is less variable and generally centers
around higher polling percentages compared to Trump's.

\hypertarget{difference-in-average-polling-percentage}{%
\subsection{Difference in Average Polling
Percentage}\label{difference-in-average-polling-percentage}}
